Experience-Biased Symbolic Planning

### 2014-05-12 AAAI_AuthorKit/paper.tex
Minor edits

From a computational perspective, the HSP heuristic must compute $g_S(a)$ for all $a\in A$ whenever a new state $S$ is generated.
This can be done by dynamic programming, iterating the recursive formula to a fixpoint in Bellman-Ford-like fashion.
In a sense, it seems we are wasting a lot of work by estimating the cost to reach every atom when we only require $g_S(G)$.
Later, we'll see how to make better use of this computation.

@@

Now we combine the ideas from HSP and E-graphs to construct a domain-independent STRIPS planner which learns from experience. In STRIPS, a consistent heuristic $h(S,S')$ is given by the HSP-max estimate $h^{HSP}_S(S')$. We can compute $h^{HSP}_S(S')$ for all $S,S'\in V^E$ in precisely $|V^E|$ runs of the dynamic programming we saw before: each run gives us $|V^E|$ of these results, yielding $(|V^E|)^2$ values in total.

Thus, our preprocessing work is roughly equivalent to that of generating $|V^E|$ states, which will be small. As before, Dijkstra's algorithm combines an inflation of these estimates with the E-graph edges to derive $h^E(S)$ for all $S\in V^E$. Note that in STRIPS, $V^E$ need not include every goal state: the minimal goal $G$ suffices as it's always the easiest to reach in the heuristic relaxation.

Upon encountering a new state $S\notin V^E$, we run dynamic programming procedure again to compute $g_S(a)$ for all atoms $a\in A$. Then it's a simple matter to compute the E-graph heuristic by

@@

We remark that, when E-graphs are involved, the values $g_S(S')$ on the right-hand side range over a variety of different states $S'$. Compare this to the formula for $h^{HSP}$, which uses only $g_S(G)$. HSP would still compute all the $g_S$-values despite using them only once, so here we get to consult the E-graph essentially for free. The only additional work is the minimization over $S'\in V^E$, which takes O($|V^E||A|$) time. A pseudocode implementation is listed in Algorithms \ref{alg:ComputeG} and \ref{alg:Search}. The parameter $\epsilon^E$ controls bias toward the E-graph, while $\epsilon$ provides an additional goal-directed bias.

\begin{algorithm}
\caption{ComputeG($S$)}
\label{alg:ComputeG}
\begin{algorithmic}
\FORALL{$a \in A$}
\IF{$a \in S$}
\STATE $g_S(a) \leftarrow 0$
\ELSE
\STATE $g_S(a) \leftarrow \infty$
\ENDIF
\ENDFOR
\REPEAT
\FORALL{$op \in O$}
\STATE $pCost \leftarrow g_S(Prec(op))$
\FORALL{$a \in Add(op)$}
\STATE $g_S(a) \leftarrow \min \left(g_S(a),~pCost + Cost(op)\right)$
\ENDFOR
\ENDFOR
\UNTIL{no further change in $g$-values}
\end{algorithmic}
\end{algorithm}

@@

ComputeG implements the dynamic programming computation of $g$-values.
Each iteration of the outermost loop permanently fixes at least one $g_S(a)$ value, so it does at most $|A|+1$ iterations.
Multiplying everything together and ignoring cardinality signs for brevity, the runtime bound is $O(\mathbf{A^2O})$.
Thus, each state generated costs $O(\mathbf{A^2O + V^EA})$ time.
The preprocessing cost is $O(\mathbf{V^EA^2O + (V^E)^2A})$, asymptotically equivalent to preemptively generating every state in $V^E$.
For comparison, standard HSP without E-graphs takes $O(\mathbf{A^2O})$ time (with essentially the same constant factors) to generate a state, and does not incur the cost of generating a state unless the A* encounters it. Provided the number of E-graph vertices is limited to $O(\mathbf{AO})$, the asymptotic cost of node generation is unchanged.

>>>>

From a computational perspective, the HSP heuristic must compute $g_S(a)$ for all $a\in A$ whenever a new state $S$ is generated.
This can be done by dynamic programming, iterating the recursive formula to a fixpoint in Bellman-Ford-like fashion.
Pseudocode is listed in Algorithm \ref{alg:ComputeG}.

\begin{algorithm}
\caption{ComputeG($S$)}
\label{alg:ComputeG}
\begin{algorithmic}
\FORALL{$a \in A$}
\IF{$a \in S$}
\STATE $g_S(a) \leftarrow 0$
\ELSE
\STATE $g_S(a) \leftarrow \infty$
\ENDIF
\ENDFOR
\REPEAT
\FORALL{$op \in O$}
\STATE $pCost \leftarrow g_S(Prec(op))$
\FORALL{$a \in Add(op)$}
\STATE $g_S(a) \leftarrow \min \left(g_S(a),~pCost + Cost(op)\right)$
\ENDFOR
\ENDFOR
\UNTIL{no further change in $g_S$-values}
\end{algorithmic}
\end{algorithm}

In a sense, a lot of work is wasted by estimating the cost to reach every atom, when we only require $g_S(G)$.
Later, we'll see how to make better use of this computation.

@@

Now we combine the ideas from HSP and E-graphs to construct a domain-independent STRIPS planner which learns from experience. In STRIPS, a consistent heuristic $h(S,S')$ is given by the HSP-max estimate $g_S(S')$. We can compute $g_S(S')$ for all $S,S'\in V^E$ in precisely $|V^E|$ runs of Algorithm \ref{alg:ComputeG}: each run provides us with $|V^E|$ of these results, yielding $(|V^E|)^2$ values in total.

As before, Dijkstra's algorithm combines an inflation of these estimates with the E-graph edges to derive $h^E(S)$ for all $S\in V^E$. Note that in STRIPS, $V^E$ need not include every goal state: the minimal goal $G$ suffices as it's always the easiest to reach in the heuristic relaxation.

Upon encountering a new state $S\notin V^E$, we run Algorithm \ref{alg:ComputeG} again to compute $g_S(a)$ for all atoms $a\in A$. Then it's a simple matter to compute the E-graph heuristic by

@@

We remark that, when E-graphs are involved, the values $g_S(S')$ on the right-hand side range over a variety of different states $S'$. Compare this to the formula for $h^{HSP}$, which uses only $g_S(G)$. HSP would still compute all the $g_S$-values despite using them only once, so here we get to consult the E-graph essentially for free. The only additional work is the minimization over $S'\in V^E$, which takes O($|V^E||A|$) time. A pseudocode implementation is listed in Algorithm \ref{alg:Search}. The parameter $\epsilon^E$ controls bias toward the E-graph, while $\epsilon$ provides an additional goal-directed bias.

@@

ComputeG($S$) implements the dynamic programming computation of $g_S$-values.
Each iteration of the outermost loop permanently fixes at least one $g_S(a)$ value, so it does at most $|A|+1$ iterations.
Multiplying everything together and ignoring cardinality signs for brevity, the runtime bound on Algorithm \ref{alg:ComputeG} is $O(\mathbf{A^2O})$.
Thus, each state generated costs $O(\mathbf{A^2O + V^EA})$ time.
The preprocessing cost is $O(\mathbf{V^EA^2O + (V^E)^2A})$, asymptotically equivalent to preemptively generating every state in $V^E$.

For comparison, standard HSP without E-graphs takes $O(\mathbf{A^2O})$ time (with essentially the same constant factors) to generate each state, and does not incur the cost of generating a state unless the A* frontier encounters it.
When HSP is augmented with small E-graphs, the same asymptotic bound continues to hold, provided that the number of E-graph vertices is limited to $O(\mathbf{AO})$. Furthermore, we retain completeness and the ability to configure suboptimality bounds by adjusting the bias parameters:
